---
title: "p8105_hw1_xy2397"
author: "Xue Yang"
date: 2018-09-16
output:
  
  html_document:
    toc: true
    toc_float: true
---



# Problem 1

## Question:
Create a data frame comprised of:

A random sample of size 10 from a uniform[0, 5] distribution
A logical vector indicating whether elements of the sample are greater than 2
A (length-10) character vector
A (length-10) factor vector
Try to take the mean of each variable in your dataframe. What works and what doesn’t? Why?

In some cases, you can explicitly convert variables from one type to another. Write a code chunk that applies the as.numeric function to the logical, character, and factor variables (please show this chunk but not the output). What happens? In a second code chunk, convert your character variable from character to factor to numeric; similarly, convert your factor variable from factor to character to numeric. What happens?


## Solution:

1. Firstly, creating the data frame

```{r creating_data_frame}
uniform_sample = runif (10, 0, 5)
logical_vector = uniform_sample > 2
character_vector = c('a', 'b', 'c', 'd', 'e', 'f', 'g', 'h', 'i', 'j')
factor_vector = factor(character_vector)
data_frame = data.frame(uniform_sample, logical_vector, character_vector, factor_vector)
```


2. Secondly, take the mean of each variable in the dataframe

* The mean of the uniform_sample is `r mean(data_frame$uniform_sample)`
* The mean of the logical_vector is `r mean(data_frame$logical_vector)`
* The mean of the character_vector is `r mean(data_frame$character_vector)`
* The mean of the factor_vector is `r mean(data_frame$factor_vector)`

From the results of mean for each variable, we can find that only vaiables uniform_sample and logical_vector have mean, that is to say, only numeric or logical can calculate the exact mean. While the variables character_vector and factor_vector are not numeric or logical (they are character or factor), as a result, R can only returning NA for them.


3. Last,  explicitly convert variables from one type to another.


*    By applies the as.numeric function to the logical, character and factor variable, we can find that all these variable changed into numeric. That is to say, for logical variable, TRUE is assigned to 1 and FALSE is assigned to 0. And for character and factor variable, they are both assigned from 1 to 10.

```{r coverting_numeric,eval = FALSE}
as.numeric(data_frame$logical_vector)
as.numeric(data_frame$character_vector)
as.numeric(data_frame$factor_vector)
```

*    By converting character variable from character to factor to numeric, we can find that the character variable finally changed into numeric from 1 to 10.

```{r converting_character_to_factor_to_numeric}

# convert character variable from character to factor
as.factor(data_frame$character_vector)

# then covert it from factor to numeric
as.numeric(as.factor(data_frame$character_vector))    
```


*    Similarly, by converting factor variable from factor to character to numeric, we can find that when we coercion covert, NAs will be introduced.
```{r coverting_factor_to-character_to_numeric}

# convert factor variable from factor to character
as.character(data_frame$factor_vector)

# then covert it from character to numeric
as.numeric(as.character(data_frame$factor_vector))
```


